Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-05-22 TEX/draft4.tex
appendix on the relation to cryo

>>>>

\section{The autocorrelation functions in cryo--EM}

The 3-D Fourier transform of an L-bandlimited 3-D volume (e.g., particle) can be expanded by  spherical harmonics:
\begin{equation} \label{eq:volume}
\hat{V}(k,\theta,\phi)  = \sum_{\ell=0}^{L} \sum_{m=-\ell}^{\ell} A_{\ell,m}(k)Y_\ell^{m}(\theta,\phi),
\end{equation}
where $(\theta,\phi)$ are two angles on the sphere, $k$ is the radial coordinate, $Y_{\ell}^m(\theta,\phi)$ is the spherical harmonic of degree $\ell$ and order $m$ and  $A_{\ell,m}(k)$ are the associated spherical harmonics coefficients, to be estimated. 
A rotation of the volume by $\omega\in SO(3)$ can be described using the Wigner D-function $D_{m,m'}^{\ell}$:
\begin{equation}
\begin{split}
(R_\omega \hat{V})(k,\theta,\phi)  &= \sum_{\ell=0}^{L} \sum_{m=-\ell}^{\ell} A_{\ell,m}(k)(R_\omega Y_\ell^{m})(\theta,\phi) \\
&= \sum_{\ell=0}^{L} \sum_{m=-\ell}^{\ell} A_{\ell,m}(k) 
\sum_{m'=-\ell}^{m=\ell}{D_{m,m'}^{\ell}(\omega)}Y_\ell^{m'}(\theta,\phi).
\end{split}
\end{equation}

By the Fourier slice theorem, each cryo--EM measurement (projection) is equivalent (through 2-D Fourier transform) to the slice of $\hat{V}$, associated with $\theta= \pi/2$, after $\hat{V}$ was rotated by $\omega\in SO(3)$. Explicitly, the Fourier transform of a projection from the viewing direction $\omega$ is related to the spherical harmonic coefficients of the object through:
\begin{equation} \label{eq:P}
\begin{split}
P_\omega(k,\phi)  = \sum_{\ell=0}^{L} \sum_{m=-\ell}^{\ell} A_{\ell,m}(k) 
\sum_{m'=-\ell}^{m=\ell}{D_{m,m'}^{\ell}(\omega)}Y_\ell^{m'}(\pi/2,\phi).
\end{split}
\end{equation}

Next, we want to relate the projections $P_\omega$ with the autocorrelation functions computed from the micrograph.
For simplicity, we ignore noise affects and assume that the projections are sufficiently separated.
The $(\Delta_x,\Delta_y)$ entry of the second-order autocorrelation of the micrograph is equal, up to a constant that depends on the number of projections, to
\begin{equation}
M_2(\Delta_x,\Delta_y) \propto \sum_{n=1}^N \sum_{x,y}P_n(x,y)P_n(x+\Delta_x,y+\Delta_y),
\end{equation}
where $P_n$ denotes the $n$th projection. By taking $n\to\infty$ and assuming uniform distribution of viewing directions, we get 
\begin{equation}
M_2(\Delta_x,\Delta_y)\propto\sum_{x,y}\int_{\omega} P_\omega(x,y)P_\omega(x+\Delta_x,y+\Delta_y)d\omega.
\end{equation}
In the same way and under the same conditions, the third moment is given by 
\begin{equation} 
M_3(\Delta_x^1,\Delta_y^1;\Delta_x^2,\Delta_y^2) \propto \sum_{x,y}\int_{\omega} P_\omega(x,y)P_\omega(x+\Delta_x^1,y+\Delta_y^1)P_\omega(x+\Delta_x^2,y+\Delta_y^2)d\omega.
\end{equation}

In order to determine the particle, by~\eqref{eq:volume} one needs to estimate an order of $L^3$ spherical harmonics coefficients.
If the pixel size is proportional to $1/L$ (to match the volume's resolution), then $M_3$ provides an order of $L^4$ equations of $P_\omega$ which depends (after coordinate transformation) linearly in the spherical harmonic coefficients through~\eqref{eq:P}.
